Regularising Lotka--Volterra interaction matrices by species grouping: optimal groupings are unpredictable

### 2022-02-22 article.rnw
Fix the response variable to be population growth, instead of offspring population size.
Fix this in both the code and the report,
including updating the results text to account for changes in the results.

The experiment used six species of \emph{Drosophila} from an Australian rainforest community: \emph{D. birchii}, \emph{D. pallidifrons}, \emph{D. pandora}, \emph{D. pseudoananassae}, \emph{D. simulans}, and \emph{D. sulfurigaster}.
The experiment took place in vials with a limited amount of medium for egg-laying and larval development; competition was for space on this medium.
A fixed number of adult pairs of one or two species were added to each vial and allowed to mate and lay eggs.
The number of adult pairs of each species emerging from these eggs---i.e.\ the size of the next generation---was the response variable.
% DBS: CRPB to fix this response variable when he fixes the model formula to reflect number of offspring.

@@

As the response variable was fecundity, I used a discrete-time version of the Lotka--Volterra competition model given by:
\[
	N_{i,t+1} = N_{i,t} r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right)
,\]
% DBS: CRPB to update the equation above after consideration of experimental response variable.
where $N_{i,t}$ is the population of species $i$ at time $t$, $r_i$ is the intrinsic growth rate of species $i$, and $\alpha_{ij}$ is the \emph{per capita} competitive effect of species $j$ on species $i$.
$N_{i,t+1}$ is the size of the population in the following time-step, i.e.\ the number of surviving offspring, the fecundity, of the previous generation, and therefore is equal to the experimentally measured response variable.
The competition coefficients were not constrained to be positive, allowing for the possibility of mutualistic effects between some species.

@@

The Lotka--Volterra model is not linear in the parameters.
With the rows of the interaction matrix ungrouped, the regression could be reparameterised to be linear in the parameters by treating $r_i \alpha_{ij}$ as a single regression parameter for each $i,j$ pair.
With the rows grouped, however, this becomes impossible since each species still has an independent growth rate.
As such, the models were fitted by the Gauss--Newton method for non-linear least squares.
%The models were fitted to the data using maximum likelihood regression techniques.
%Fitting it using precisely a Lotka--Volterra model prevented the use of a link function, rendering most generalised regression techniques infeasible.
I assumed a normal error distribution of the response variable and to prevent heteroskedasticity and improve normality, I transformed total fecundity to \emph{per capita} fecundity, which corresponds to the following model:
\[
	\frac{N_{i,t+1}}{N_{i,t}} = r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right)
.\]
% DBS: CRPB to alter equation above consist with changes to previous equations.

@@

In all, no pair of species can confidently be said to be coclassified, in the row or column groupings; no probability of coclassification exceeded \SI{60}{\percent}.
Some pairs of species were far more likely to be coclassified than others, however, suggesting some structure to the best groupings.
% DBS: As I have noted before, I'd like to know what the null expectation is for this before claiming confidence of being co-classified or not being co-classified. Firstly, I think that the random expectation is much lower than one might otherwise suspect. (For three species, it is already 2/5; for four species, it is 5/15.) Second, the conventional "ecological" null is actually that the probability is 0, in which case it doesn't take much co-classification probability to see the glass as half full and argue that grouping is a much more parsimonious approach than the contrary.

@@

A high likelihood of coclassification in the row groupings was typically transitive: e.g.\ \emph{D. birchii}, \emph{D. pallidifrons} and \emph{D. sulfurigaster} each had a relatively high probability of being coclassified with one another.
More intransitivity was observed in the column groupings: \emph{D. simulans} had a relatively high probability of being coclassified with \emph{D. sulfurigaster}, and \emph{D. sulfurigaster} had a substantial probability of being coclassified with \emph{D. pallidifrons}, but \emph{D. simulans} had a much lower probability of being coclassified with \emph{D. pallidifrons}.

A high likelihood of coclassification was associated with phylogeny in one instance, but by no means consistently.
The phylogeny divided the six species into three pairs of most closely related species (\cref{fig:phylogeny}).
The pair of \emph{D. pallidifrons} and \emph{D. sulfurigaster}, which both belonged to a different subgenus than the other four species, were also the only pair of species to be coclassified with relatively high probability in both the row and column groupings.
In contrast, both of the other most closely related pairs were coclassified with very low probabilities in both the row and column groupings.
Furthermore, both row and column groupings gave high probabilities of coclassification to pairs of species split across the two subgenera in the study---the row groupings frequently coclassified \emph{D. sulfurigaster}of the Drosophila subgenus and \emph{D. simulans} of the Sophophora subgenus, while the column groupings frequently coclassified both \emph{D. pallidifrons} and \emph{D. sulfurigaster} of the Drosophila subgenus with \emph{D. birchii} of the Sophophora subgenus.

>>>>

The experiment used six species of \emph{Drosophila} from an Australian rainforest community: \emph{D. birchii}, \emph{D. pallidifrons}, \emph{D. pandora}, \emph{D. pseudoananassae}, \emph{D. simulans}, and \emph{D. sulfurigaster}.
The experiment took place in vials with a limited amount of medium for egg-laying and larval development; competition was for space on this medium.
A fixed number of adult pairs of one or two species were added to each vial and allowed to mate and lay eggs.
The number of adult pairs of each species emerging from these eggs was counted---this was the size of the next generation.

@@

The data provided the change population size from one generation to the next, so I used a discrete-time version of the Lotka--Volterra competition model given by:
\[
	N_{i,t+1} - N_{i,t} = N_{i,t} r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right)
,\]
where $N_{i,t}$ is the population of species $i$ at time $t$ (the size of the parent generation), $N_{i,t+1}$ is the population at time $t+1$ (the size of the offspring generation), $r_i$ is the intrinsic growth rate of species $i$, and $\alpha_{ij}$ is the \emph{per capita} competitive effect of species $j$ on species $i$.
The competition coefficients were not constrained to be positive, allowing for the possibility of mutualistic effects between some species.

@@

The Lotka--Volterra model is not linear in the parameters.
With the rows of the interaction matrix ungrouped, the regression could be reparameterised to be linear in the parameters by treating $r_i \alpha_{ij}$ as a single regression parameter for each $i,j$ pair.
With the rows grouped, however, this becomes impossible since each species still has an independent growth rate.
As such, the models were fitted by the Gauss--Newton method for non-linear least squares.
%The models were fitted to the data using maximum likelihood regression techniques.
%Fitting it using precisely a Lotka--Volterra model prevented the use of a link function, rendering most generalised regression techniques infeasible.
I assumed a normal error distribution of the response variable and to prevent heteroskedasticity and improve normality, I transformed total population growth to \emph{per capita} population growth, which corresponds to the following model:
\[
	\frac{N_{i,t+1} - N_{i,t}}{N_{i,t}} = r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right)
.\]

@@

In all, no pair of species can confidently be said to be coclassified, in the row or column groupings; no probability of coclassification exceeded \SI{75}{\percent}.
Some pairs of species were far more likely to be coclassified than others, however, suggesting some structure to the best groupings.
% DBS: As I have noted before, I'd like to know what the null expectation is for this before claiming confidence of being co-classified or not being co-classified. Firstly, I think that the random expectation is much lower than one might otherwise suspect. (For three species, it is already 2/5; for four species, it is 5/15.) Second, the conventional "ecological" null is actually that the probability is 0, in which case it doesn't take much co-classification probability to see the glass as half full and argue that grouping is a much more parsimonious approach than the contrary.

@@

A relatively high likelihood of coclassification in the row groupings was typically transitive: e.g.\ \emph{D. birchii}, \emph{D. pallidifrons} and \emph{D. sulfurigaster} each had a relatively high probability of being coclassified with one another.
More intransitivity was observed in the column groupings: \emph{D. pallidifrons} had a relatively high probability of being coclassified with \emph{D. sulfurigaster}, and \emph{D. sulfurigaster} had a substantial probability of being coclassified with \emph{D. simulans}, but \emph{D. pallidifrons} had a much lower probability of being coclassified with \emph{D. simulans}.

A high likelihood of coclassification was associated with phylogeny in one instance, but by no means consistently.
The phylogeny divided the six species into three pairs of most closely related species (\cref{fig:phylogeny}).
The pair of \emph{D. pallidifrons} and \emph{D. sulfurigaster}, which both belonged to a different subgenus than the other four species, were also the only pair of species to be coclassified with relatively high probability in both the row and column groupings.
In contrast, both of the other most closely related pairs were coclassified with very low probabilities in both the row and column groupings.
